Testing Charts: viewers' perceptual accuracy in surveys

### 2023-07-27 SDSS-2023/testing-charts-accuracy.tex
a bunch of NDT comments addressed

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{lscape}
\author[1]{Kiegan Rice\thanks{Corresponding author. Email: rice-kiegan@norc.org}}
\author[2]{Heike Hofmann\footnote{Email: hofmann@iastate.edu}}
\author[1]{Nola du Toit\footnote{Email: dutoit-nola@norc.org}}
\author[1]{Edward Mulrow\footnote{Email: mulrow-edward@norc.org}}
\affil[1]{NORC at the University of Chicago}
\affil[2]{Department of Statistics, Iowa State University}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Testing Charts: viewers' perceptual accuracy in surveys},
  pdfkeywords={list, keywords, here, 3-6, are, fine},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle
\begin{abstract}
The use of visuals is a key component in scientific communication, and
decisions about the design of a data visualization should be informed by
what design elements best support the audience's ability to perceive and
understand the components of the data visualization. We build on the
foundations of Cleveland and McGill's work in graphical perception,
employing a large, nationally-representative, probability-based panel of
survey respondents to test perception in statistical charts. Our
findings provide actionable guidance for data visualization
practitioners to employ in their work.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

What do viewers see when we show them a data chart? A data chart -- at
its core -- maps values to graphical elements: quantitative elements are
represented as measurable features, such as position, size, or shade.
\hh{do we need to also go into qualitative variables?} Modern data
visualizations are much more than a simple, objective mapping of values
to a plane; they contain contextual and design elements, and are often
structured to support the viewer in understanding a particular view of a
set of data or specific pattern underlying the values. The design of a
data visualization impacts a viewer's ability to achieve that
understanding; a poorly designed data visualization may leave viewers
struggling to understand the content or context, or make it difficult to
complete accurate and useful comparisons of values across groups or time
points. More broadly, the design of a data visualization can change how
viewers interact with the chart.

\ndt{define structural}

\ndt{The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding of chart.}

Cleveland and McGill (1984) observed as such, and in their seminal study
defined the better visual among a pair as the one that allows viewers to
make more accurate comparisons. Based on mappings of quantitative
variables to different graphical elements
\ndt{example of element to explain it XXX}, Cleveland and McGill's study
resulted in a ranking of perceptual tasks \ndt{give example of task}
from most accurate to least accurate, which was then extended by
Mackinlay (1986) to a theoretical framework ranking tasks' order along
their ordinal and nominal scales.

@@

Here, we seek to answer whether it is possible to reproduce some of the
previous findings in the context of a survey with a large,
nationally-representative set of respondents, and within that context we
focus on the following research questions:

@@

We employ a probability-based survey panel and run a series of
perception tests with nationally-representative samples of respondents
from that panel. The advantage of using a probability-based approach is
two-fold. First, we have access to a large sample of survey participants
and thus have greater power in making inference about graphical
perceptional abilities. Second, the sample is representative of the
general adult public in the U.S., and this allows us to test whether
prior results from convenience samples hold with a nationally
representative sample and whether there are differences in those results
across demographic subgroups.

\ndt{I think the next paragraph can be moved above the RQs. Use this paragraph to set up the concepts, define constructs, then explain the study, and then note the RQs.}

Specifically, we present viewers with structural variations of bar
charts and ask them to answer questions comparing the size of elements
within those charts. In this work, we present the series of tests we
completed and the resulting findings. The remainder of the paper is
organized as follows: first, we describe the design of the visual
stimuli used in the perception tests. We then describe the population of
study respondents and obtained survey sample. Subsequently, we share a
summary of the resulting survey responses across each of the tests,
including analyses on accuracy of responses and response behavior.
Finally, we discuss implications of this work and next steps.

@@

\ndt{
I think this paragraph should go first in this section.
1. the survey format
2. the type of data
3. The stimulus
3. how we set up the stimulus and why
4. the tasks and alighed v unaligned
4. qtext
5. fig 1}

@@

We employ comparisons at the JND in our tasks in order to maximize our
ability to identify the impact of design changes on viewer accuracy and
behavior. Asking perception tasks in a survey differs from the
controlled environment of a cognitive lab, where these kind of questions
may usually be assessed. Rather than asking the same (or similar) type
of question with varied signal strength dozens or hundreds of times, we
are limited to only a few questions at a time.
\ndt{note why we are limited to only a few questions at a time - that is not clear.}
With a small set of tasks, we need to present tasks that are
perceptually hard, and thus ask questions about stimuli that are close
to our perceptual threshold. Therefore, we focus on questions which vary
the structure of the presented image, but ask viewers to compare the
same underlying data values across those varied images.

We ask participants to determine which of two just-noticeably different
marked pieces \ndt{is pieces the same as tiles below?} is larger within
each chart, and focus on three main sets of structural variation in the
design of that chart. First, we vary the alignment of the pieces in
question. \ndt{perhaps a bulleted list would work better here} Viewers
are presented with two marked pieces in a chart that do not share a
common baseline, then two pieces that do share a common baseline.
\ndt{"we refer to this as aligned and unaligned"} Second, we vary the
orientation of the chart -- we rotate the vertical stacked bar chart,
and present a horizontally oriented version of the same chart, with
identically sized marked pieces. Finally, we change the aspect ratio of
the chart and present a wider version of the horizontally oriented chart
which has longer, but thinner, marked pieces.

\hh{Not sure where should put this, parking it here for right now:}
\hh{expectations - start}

What we call `aligned' and `unaligned', here, is similar to Cleveland
and McGill's positions along aligned and unaligned scales, but with some
modifications. Figure~\ref{fig-tasks-explain} gives an overview of the
comparisons of tasks 1 through 3 and the closest corresponding tasks in
Cleveland and McGill.

In this paper, both `aligned' and `unaligned' bars share the same axis.

Aligned tiles are additionally anchored in the same position along the
axis, i.e.~the difference between their sizes can be reduced to a
positional assessment, which in the absence of three-dimensional cues
seems to be the general strategy viewers use when evaluating barcharts
(Zacks et al. 1998).\\
Unaligned tiles do not share this anchor of position, i.e.~the
difference between their sizes needs at least two comparisons of
positions. However, unaligned bars in a stacked barchart are not just
assessments of the lengths of bars -- the other tiles in the chart
provide a reference frame of the shared axis, which \emph{should} help
with an assessment of the tiles' sizes. As shown in the sketch of
Figure~\ref{fig-tasks-explain}, this framing leads to a stimulus that is
between Cleveland and McGill's task of `position along the same but
unaligned axis' and an assessment of length.

We would expect that comparing unaligned tiles is a harder task (with
correspondingly lower levels of accuracy) than a comparison of aligned
tiles, with the framing given by the other tiles in the same column
mitigating some of this difficulty.
\hh{In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors}
(Talbot, Setlur, and Anand 2014).

\caption{\label{fig-tasks-explain}Comparisons made in charts within the
Cleveland and McGill ranking}

\ndt{"The survey format guides the design and structure of the questions asked...." Why is this so?}

Our use of a survey format guides the format and design of the questions
asked and how they are presented to respondents. First, participant
instructions must be delivered in a very short and easily understandable
format, because participants cannot ask clarifying questions about the
task as they might be able to in a cognitive lab setting. Second, we
want to utilize content within the data visualization image which is not
socially or politically charged for the average U.S. adult; this risks
participants reacting to the subject matter within the chart rather than
focusing on the presented task. \ndt{stimulus?} For this reason, we
utilize data on living arrangements of older U.S. adults -- a topic
which most U.S. adults will have some familiarity with, but is not
inherently polarizing. Finally, to prevent viewers from being exposed to
slight variations of the same stimulus in a row (and risk unforeseen
order effects or respondents using prior questions to inform their
responses), we either split a survey sample in two and show each
subsample a distinct structural version of the chart or test variations
of a chart across distinct rounds of the survey.

\ndt{This should go above fig 1 to explain the set up for fig 1}

@@

Task 1 was presented to viewers in two rounds (Rounds 1-2), with four
color scheme variations each presented to 50\% of respondents in each
round. There were no significant differences in respondent accuracy or
behavior across color schemes, and we combine them here into one set of
stimulus responses to focus on the comparison of structural differences.
\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? }
\hh{we could do that, but I'm not sure that we need to mention color at all.}

@@

The images utilized in the third task again represent the same data as
in the first two tasks, and the overall image has dimensions which are
identical to the images in the first two tasks, displaying at the same
size during the survey.
\hh{The length of bars is increased to fit the new dimensions. This increases the difference in the length of bars from previously 8 pixels to 13 pixels. The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant. }
\ndt{agree to drop note on color. Can perhaps just say we used CBlind  friendly}

\hypertarget{task-3b-floating-bars}{%
\subsubsection{Task 3b: Floating bars}\label{task-3b-floating-bars}}

\hh{
In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}

@@

Participants were recruited as part of NORC's AmeriSpeak panel, which
utilizes a probability-based sampling methodology and samples U.S.
households from NORC's National Sample Frame that provides coverage of
over 97\% of U.S. households. The current panel size is 54,001 panel
members aged 13 and over residing in over 43,000 households (Dennis
2019, updated 2022). \textbackslash ejm\{46,565 panelists are CAWI; they
are 93.3\% of the panel and 93.6\% of the total panel weights.\} Each
test was conducted using the AmeriSpeak Omnibus survey, which runs
biweekly and samples around 1,000 U.S. adults to answer questions on a
variety of topics. Our tests require visual inspection of an image; for
this reason, our survey questions were only presented to web-based
panelists and not panelists who respond via phone interviews.

@@

Paragraph on how survey weights are derived? -- \textbf{Question for Ed
-- is this something AmeriSpeak can provide boilerplate language for?}

All calculations in this paper are done in R (R Core Team 2022), and
weights are applied in analyses using the \texttt{survey} package
(Lumley 2004) version 4.0 (Lumley 2020) based on Lumley (2010).

When combining responses from different surveys, weights
\ejm{are rescaled} with respect to the total population, so that their
order after combining still reflects their relative importance.

We \emph{combine} (rather than cumulate) a set of \(\ell\) surveys
\(S_1\), \(S_2\), \ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as
described in O'Muircheartaigh and Pedlow (2002), by multiplying weights
in \(S_i\) by \(\lambda_i\), for \(1 \le i \le \ell\).
\(\lambda_i \in [0,1]\) with \(\sum_i \lambda_i = 1\) is given as \[
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\] where \(n_i\) is the nominal sample of survey \(S_i\) and \(d_i\) are
the design effects for the estimators. Here, \(d_i\) are estimated as \[
d_i = 1 + CV(w \in S_i)^2
\] where \(CV\) is the coefficient of variation of the weights \(w\)
within each sample, and is estimated as in Kish (1965) : \[ 
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\]

@@

\hypertarget{tbl-rounds}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{\label{tbl-rounds}Survey rounds: dates, number of participants
(nominal sample size), effective sample size, sum of weights, and
factors for the combining surveys as discussed above.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Round 1 & April 2022 & 933 & 521.1 & 934.9 & \(\lambda_1\) = 0.343 \\
Round 2 & May 2022 & 953 & 485.7 & 953.4 & \(\lambda_2\) = 0.320 \\
Round 3 & June 2022 & 921 & 513.5 & 923.1 & \(\lambda_3\) = 0.338 \\
\textbf{Total} & --- & 2807 & 1520.8 & & \\
\end{longtable}

@@

\hh{The actionable insights are:}

>>>>

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{lscape}
\author[1]{Kiegan Rice\thanks{Corresponding author. Email: rice-kiegan@norc.org}}
\author[2]{Heike Hofmann\footnote{Email: hofmann@iastate.edu}}
\author[1]{Nola du Toit\footnote{Email: dutoit-nola@norc.org}}
\author[1]{Edward Mulrow\footnote{Email: mulrow-edward@norc.org}}
\affil[1]{NORC at the University of Chicago}
\affil[2]{Department of Statistics, Iowa State University}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Testing Charts: viewers' perceptual accuracy in surveys},
  pdfkeywords={list, keywords, here, 3-6, are, fine},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{document}
\maketitle
\begin{abstract}
The use of visuals is a key component in scientific communication.
Decisions about the design of a data visualization should be informed by
what design elements best support the audience's ability to perceive and
understand the components of the data visualization. We build on the
foundations of Cleveland and McGill's work in graphical perception,
employing a large, nationally-representative, probability-based panel of
survey respondents to test perception in statistical charts. Our
findings provide actionable guidance for data visualization
practitioners to employ in their work.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, frame hidden]}{\end{tcolorbox}}\fi

@@

What do viewers see when we show them a data chart? A data chart -- at
its core -- maps values to graphical elements: quantitative elements are
represented as measurable features, such as position, size, or shade.
Modern data visualizations are much more than a simple, objective
mapping of values to a plane; they contain contextual and design
elements, and are often structured to support the viewer in
understanding a particular view of a set of data or specific pattern
underlying the values. Structural choices, such as the choice of a
graphical element (e.g., bar or pie), orientation of the chart, and
sizing and relative positioning of elements within the chart, are
determined by the data visualization practitioner. The design of a data
visualization impacts a viewer's ability to achieve that understanding;
a poorly designed data visualization may leave viewers struggling to
understand the content or context, or make it difficult to complete
accurate and useful comparisons of values across groups or time points.
More broadly, the design of a data visualization can change how viewers
interact with the chart.

The viewer's employment of comparisons of the components is a crucial
step in the process of interacting with and understanding of chart.
Cleveland and McGill (1984) observed as such, and in their seminal study
defined the better visual among a pair as the one that allows viewers to
make more accurate comparisons. Based on mappings of quantitative
variables to different graphical elements, Cleveland and McGill's study
resulted in a ranking of perceptual tasks (such as ranking the larger of
two lengths or areas) from most accurate to least accurate, which was
then extended by Mackinlay (1986) to a theoretical framework ranking
tasks' order along their ordinal and nominal scales.

@@

Here, we seek to answer whether it is possible to reproduce some of the
previous findings in the context of a survey with a large,
nationally-representative set of respondents. Specifically, we present
viewers with structural variations of bar charts and ask them to answer
questions comparing the size of elements within those charts. We employ
a probability-based survey panel and run a series of perception tests
with nationally-representative samples of respondents from that panel.
The advantage of using a probability-based approach is two-fold. First,
we have access to a large sample of survey participants and thus have
greater power in making inference about graphical perceptional
abilities. Second, the sample is representative of the general adult
public in the U.S., and this allows us to test whether prior results
from convenience samples hold with a nationally representative sample
and whether there are differences in those results across demographic
subgroups. Within that context we focus on the following research
questions:

@@

In this work, we present the series of tests we completed and the
resulting findings. The remainder of the paper is organized as follows:
first, we describe the design of the visual stimuli used in the
perception tests. We then describe the population of study respondents
and obtained survey sample. Subsequently, we share a summary of the
resulting survey responses across each of the tests, including analyses
on accuracy of responses and response behavior. Finally, we discuss
implications of this work and next steps.

@@

Our use of a survey format guides the format and design of the questions
asked and how they are presented to respondents. First, participant
instructions must be delivered in a very short and easily understandable
format, because participants cannot ask clarifying questions about the
task as they might be able to in a cognitive lab setting. Our tasks are
presented as a short series of questions within a larger survey on a
variety of topics. Second, we want to utilize content within the
stimulus which is not socially or politically charged for the average
U.S. adult; this risks participants reacting to the subject matter
within the chart rather than focusing on the task. For this reason, we
utilize data on living arrangements of older U.S. adults -- a topic
which most U.S. adults will have some familiarity with, but is not
inherently polarizing. Finally, to prevent viewers from being exposed to
slight variations of the same stimulus in a row (and risk unforeseen
order effects or respondents using prior questions to inform their
responses), we either split a survey sample in two and show each
subsample a distinct structural version of the chart or test variations
of a chart across distinct rounds of the survey.

@@

We employ comparisons at the JND in our tasks in order to maximize our
ability to identify the impact of design changes on viewer accuracy and
behavior. Asking perception tasks in a survey differs from the
controlled environment of a cognitive lab, where these kind of questions
may usually be assessed. Rather than asking the same (or similar) type
of question with varied signal strength dozens or hundreds of times, we
are limited to only a few questions at a time. With a small set of
tasks, we need to present tasks that are perceptually hard, and thus ask
questions about stimuli that are close to our perceptual threshold.
Therefore, we focus on questions which vary the structure of the
presented image, but ask viewers to compare the same underlying data
values across those varied images.

We ask participants to determine which of two just-noticeably different
marked pieces (tiles) is larger within each chart, and focus on three
main sets of structural variation in the design of that chart:

\begin{itemize}
\item
  First, we vary the alignment of the pieces in question. Viewers are
  presented with two marked pieces in a chart that do not share a common
  baseline (aligned pieces), then two pieces that do share a common
  baseline (unaligned pieces).
\item
  Second, we vary the orientation of the chart -- we rotate the vertical
  stacked bar chart, and present a horizontally oriented version of the
  same chart, with identically sized marked pieces.
\item
  Finally, we change the aspect ratio of the chart and present a wider
  version of the horizontally oriented chart which has longer, but
  thinner, marked pieces.
\end{itemize}

@@

@@

The images utilized in the third task, \autoref{fig-tasks}(c,f), again
represent the same data as in the first two tasks, and the overall image
has dimensions which are identical to the images in the first two tasks,
displaying at the same size during the survey. The length of bars is
increased to fit the new dimensions. This increases the difference in
the length of bars from previously 8 pixels to 13 pixels. The widths of
the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the
overall area of the tiles approximately constant.

\caption{\label{fig-tasks-explain}Comparisons made in charts within the
Cleveland and McGill ranking (left) and their corresponding
representations in our study tasks (right).}

\hypertarget{alignment}{%
\paragraph{Alignment}\label{alignment}}

What we call `aligned' and `unaligned', here, is similar to Cleveland
and McGill's positions along aligned and unaligned scales, but with some
modifications. Figure~\ref{fig-tasks-explain} gives an overview of the
comparisons of tasks 1 through 3 and the closest corresponding tasks in
Cleveland and McGill. In this paper, both `aligned' and `unaligned' bars
share the same axis.

Aligned tiles are additionally anchored in the same position along the
axis, i.e.~the difference between their sizes can be reduced to a
positional assessment, which in the absence of three-dimensional cues
seems to be the general strategy viewers use when evaluating barcharts
(Zacks et al. 1998). Unaligned tiles do not share this anchor of
position, i.e.~the difference between their sizes needs at least two
comparisons of positions. However, unaligned bars in a stacked barchart
are not just assessments of the lengths of bars -- the other tiles in
the chart provide a reference frame of the shared axis, which
\emph{should} help with an assessment of the tiles' sizes. As shown in
the sketch of Figure~\ref{fig-tasks-explain}, this framing leads to a
stimulus that is between Cleveland and McGill's task of `position along
the same but unaligned axis' and an assessment of length.

We would expect that comparing unaligned tiles is a harder task (with
correspondingly lower levels of accuracy) than a comparison of aligned
tiles, with the framing given by the other tiles in the same column
mitigating some of this difficulty. In particular, the additional
context of the other tiles should \emph{help} rather than \emph{hurt}
the comparison, unlike other situations, where additional information
besides the pieces involved in the comparison were found to be
distractors (Talbot, Setlur, and Anand 2014).

@@

Participants were recruited as part of NORC's AmeriSpeak
panel\footnote{NORC is a non-partisan organization that collects data
  for government and non-government clients}, which utilizes a
probability-based sampling methodology and samples U.S. households from
NORC's National Sample Frame that provides coverage of over 97\% of U.S.
households. The current panel size is 54,001 panel members aged 13 and
over residing in over 43,000 households (Dennis 2019, updated 2022).
Each test was conducted using the AmeriSpeak Omnibus survey, which runs
biweekly and samples around 1,000 U.S. adults to answer questions on a
variety of topics. Our tests require visual inspection of an image; for
this reason, our survey questions were only presented to web-based
panelists and not panelists who respond via phone interviews. The panel
has 45,565 web-based panelists, representing 93.3\% of the total panel
and 96.3\% of the total panel weights.

@@

The sample used for each Omnibus study is selected from the AmeriSpeak
Panel using 48 sampling strata, split by age, race/Hispanic ethnicity,
education, and gender, with the size of each sample per sampling stratum
determined by the population distribution for each stratum. In addition,
sample selection takes into account expected differential survey
completion rates by demographic groups in order to achieve a
representative sample of the target population. Collected data are
weighted to the latest Current Population Survey (CPS) benchmarks from
the U.S. Census Bureau. They are balanced by gender, age, education,
race/ethnicity, and geographic region. All calculations in this paper
are done in R (R Core Team 2022), and weights are applied in analyses
using the \texttt{survey} package (Lumley 2004) version 4.0 (Lumley
2020) based on Lumley (2010).

When combining responses from different surveys, weights are rescaled
with respect to the total population, so that their order after
combining still reflects their relative importance.

We \emph{combine} (rather than cumulate) a set of \(\ell\) surveys
\(S_1\), \(S_2\), \ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as
described in O'Muircheartaigh and Pedlow (2002), by multiplying weights
in \(S_i\) by \(\lambda_i\), for \(1 \le i \le \ell\).
\(\lambda_i \in [0,1]\) with \(\sum_i \lambda_i = 1\) is given as

\[
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\]

where \(n_i\) is the nominal sample of survey \(S_i\) and \(d_i\) are
the design effects for the estimators. Here, \(d_i\) are estimated as

\[
d_i = 1 + CV(w \in S_i)^2
\]

where \(CV\) is the coefficient of variation of the weights \(w\) within
each sample, and is estimated as in Kish (1965) :

\[ 
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\]

@@

\hypertarget{tbl-rounds}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{\label{tbl-rounds}Survey rounds: dates, number of participants
(nominal sample size), effective sample size, sum of weights, and
factors for the combining surveys as discussed above.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule()
\endhead
Round 1 & April 2022 & 933 & 521.1 & 934.9 & \(\lambda_1\) = 0.343 \\
Round 2 & May 2022 & 953 & 485.7 & 953.4 & \(\lambda_2\) = 0.320 \\
Round 3 & June 2022 & 921 & 513.5 & 923.1 & \(\lambda_3\) = 0.338 \\
\textbf{Total} & --- & 2807 & 1520.8 & & \\
\bottomrule()
\end{longtable}

@@
